%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[sigplan,10pt]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
  
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[sigplan]{acmart}\settopmatter{}

%% Conference information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmConference[POPL 2018]{}{January 07--13, 2018}{Los Angeles, CA, USA}
%\acmYear{2017}
%\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
%\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
%\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2017}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{acm-reference-format}
%% Citation style
\citestyle{acmnumeric}     %% For numeric citations


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from traditional SIGPLAN
%% proceedings format to PACMPL format must update the
%% '\documentclass' and topmatter commands above; see
%% 'acmart-pacmpl-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{listings}

\input{definitions.tex}

\begin{document}

%% Title information
\title[ott-ifc]{\ottifc: Automatically Generating Information-Flow Control Mechanisms}


%% Author information
\author{Andrew Bedford}
\orcid{0000-0003-3101-4272}             %% \orcid is optional
\affiliation{
  \institution{Laval University}            %% \institution is required
  \state{Quebec}
  \country{Canada}                    %% \country is recommended
}
\email{andrew.bedford.1@ulaval.ca}          %% \email is recommended


%% Abstract
\begin{abstract}
We present \ottifc\, a tool that can automatically generate information-flow control mechanisms from programming language specifications (i.e., syntax and semantics).
\end{abstract}

\maketitle

%Problem and Motivation: Clearly state the problem being addressed and explain the reasons for seeking a solution to this problem.
\section{Problem and Motivation}
Modern operating systems rely mostly on access-control mechanisms to protect users information. However, access control mechanisms are insufficient as they cannot regulate the propagation of information once it has been released for processing. To address this issue, a new research trend called \emph{language-based information-flow security}~\cite{DBLP:journals/jsac/SabelfeldM03} has emerged. The idea being to use techniques from programming languages, such as program analysis, monitoring, rewriting and type checking, to enforce information-flow policies (e.g., information from a private file should not saved in a public file). Mechanisms that enforce information-flow policies (e.g., \cite{DBLP:journals/compsec/BedfordCDKT17}) are called \emph{information-flow control mechanisms}. 

Due to the numerous ways through which information may flow in a program, developping sound information-flow control mechanisms can be a challenging and error-prone task, particularly when dealing with complex programming languages.

%This is usually done by associating sensitive information with a label, which is then propagated whenever the information is used; a process called tainting. 

%Background and Related Work: Describe the specialized (but pertinent) background necessary to appreciate the work. Include references to the literature where appropriate, and briefly explain where your work departs from that done by others.
\section{Background and Related Work}
Most information-flow control mechanisms seek to enforce a policy called \emph{non-interference}~\cite{DBLP:conf/sp/GoguenM82a}. It essentially states that private information may not interfere with the publicly observable behavior of a program. To enforce non-interference, a mechanism must take into account two types of information flows: \emph{explicit flow} and \emph{implicit flow}~\cite{DBLP:journals/cacm/Denning76}. An insecure explicit information flow (Listing~\ref{listing:explicit-flow}) occurs when a private information flows directly into public information. An insecure implicit information flow (Listing~\ref{listing:implicit-flow}) occurs when private information influences public information through the control-flow of the application.

\begin{lstlisting}[captionpos=b, caption=Insecure explicit flow, label=listing:explicit-flow]
  public := private
\end{lstlisting}

\begin{lstlisting}[captionpos=b, caption=Insecure implicit flow, label=listing:implicit-flow]
  if (private > 0) then
    public := 0
  else
    public := 1
  end
\end{lstlisting}

Techniques for tracking and controlling explicit and implicit flows at the language level are well known~\cite{DBLP:journals/jcs/VolpanoIS96, DBLP:conf/csfw/RussoS10}, and are used in this work. 

%The generated mechanisms enforce non-interference and take into account explicit and implicit flows. 

%Approach and Uniqueness: Describe your approach in attacking the problem and clearly state how your approach is novel.
\section{Approach and Uniqueness}
We have developped a tool called \ottifc\ that can generate basic information-flow control mechanisms from programming language specifications (i.e., syntax and semantics). To the best of our knowledge, we are the first ones to achieve this; they are usually designed and implemented from the ground-up by a human.

As the name implies, \ottifc\ uses Ott~\cite{DBLP:journals/jfp/SewellNOPRSS10} as its input/output language. Ott is a tool that takes as input a programming language's syntax and semantics, written in a concise ASCII notation that ressembles what one would write in informal mathematics, and generates LaTeX, Coq or Isabelle/HOL versions of the definition.

\ottifc\ makes two assumptions about the language's specification: (1) that its syntax be composed of expressions, which may only read the memory, and commands, which may read or write the memory; and (2) that its semantics program configurations be of the form $\langle command, memory\rangle$. 

To illustrate our approach, consider the simple imperative language of Listing~\ref{listing:input-syntax}.
\begin{lstlisting}[label=listing:input-syntax,captionpos=b,caption=Ott syntax of a simple imperative language]
  arith_expr, a ::= x | n | a1 + a2 | a1 * a2 
  bool_expr, b ::= true | false | a1 < a2
  commands, c ::= skip | x := a | c1 ; c2 | 
                  if b then c1 else c2 end | 
                  while b do c end   
\end{lstlisting}


We start by inserting in all the rules a typing environment \lstinline{E} which maps variables to their level of information, and a program counter \lstinline{pc} is used to keep track of the level of information that influenced the control-flow.

\newpage
To prevent explicit flows, we identify the semantics rules that may modify the memory \lstinline{m} (e.g., Listing~\ref{listing:input-semantics-assign}). 

\begin{lstlisting}[label=listing:input-semantics-assign,captionpos=b,caption={Ott big-step semantics of the assign command}]
  <a, m> || <n, m>
  ---------------------------------
  <x := a, m> || <skip, m[x |-> n]>
\end{lstlisting}
In each of those rules, we insert a condition to ensure that we are not storing private information in a public variable.

%\begin{lstlisting}[label=listing:input-semantics-assign,captionpos=b,caption={Ott big-step semantics of the assign command}]
%  <a, m> || <n, m>
%  E |- a : ta
%  E |- x : tx
%  ta <= tx
%  ---------------------------------
%  <x:=a, m, E, pc> || <skip, m[x |-> n], E[x |-> ta], pc>
%\end{lstlisting}

To prevent implicit flows, we identify commands that may influence the control-flow of the application. That is, commands for which a program configuration may lead to two different program configurations (e.g., Listing~\ref{listing:input-semantics-if}).

\begin{lstlisting}[label=listing:input-semantics-if, captionpos=b,caption={Ott big-step semantics of the if command}]
  <b, m> || <true, m>
  <c1, m> || <skip, m1>
  --------------------------------------------
  <if b then c1 else c2 end, m> || <skip, m1>
  
  <b, m> || <false, m>
  <c2, m> || <skip, m2>
  --------------------------------------------
  <if b then c1 else c2 end, m> || <skip, m2>
\end{lstlisting}
We then update to the program counter \lstinline{pc} with the level of the expressions that are present in the rule (only \lstinline{b} in this case).
\begin{lstlisting}[label=listing:output-semantics-if, captionpos=b,caption={Instrumented semantics of the if command}]
  <b, m> || <true, m>
  E |- b : tb
  <c1, m, E, pc |_| tb> || <skip, m1, E1, pc1>
  --------------------------------------------
  <if b then c1 else c2 end, m, E, pc> || <skip, m1, E1, pc>
  
  <b, m> || <false, m>
  E |- b : tb
  <c2, m, E, pc |_| tb> || <skip, m2, E2, pc2>
  --------------------------------------------
  <if b then c1 else c2 end, m, E, pc> || <skip, m2, E2, pc>
\end{lstlisting}

\section{Current Status and Future Work}
We have implemented a prototype of our algorithm and validated that it works on two  imperative languages: one defined using small-step semantics and the other using big-step semantics. We have also begun to draft a soundness proof, that is, a proof showing that the generated mechanisms enforce non-interference.

Before our tool can be of real use to most researchers, much work remains to be done.

\paragraph{Language Support} Our assumptions restricts the types of languages that can be used in \ottifc. For example, most functional languages would not be supported because, in those languages, functions can be expressions. We are currently in the process of building a repository of formalized languages so that we can test and extend our approach to a wider range of languages.

\paragraph{Parametrization} For the moment, \ottifc\ only generates one type of information-flow control mechanism: a monitor. We plan on parametrizing our tool so that users can choose the type of mechanism to generate (e.g., type system, monitor) and choose some of its features (e.g., flow-sensitivity, termination-sensitivity, progress-sensitivity).

\paragraph{Generating Formal Proofs} We expect that some users will use the mechanisms generated by \ottifc\ as a foundation to build better and more precise mechanisms. One of the most grueling task when building an information-flow control mechanism is to prove its soundness. In order to help those users, we plan on generating a skeleton of the proof in Coq or Isabelle/HOL (both languages are supported by Ott).

\paragraph{Verifying Existing Mechanisms} The same rules that \ottifc\ uses to generate sound mechanisms could be used to verify the soundness of existing mechanisms and identify potential errors.

%% Acknowledgments
\begin{acks}
We would like to thank Josée Desharnais and Nadia Tawbi for their support and the anonymous reviewers for their comments.
\end{acks}

%% Bibliography
\bibliography{references}

\end{document}
